\documentclass{article}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{float}
\usepackage{verbatim}

\linespread{1.3}
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}
\setcounter{secnumdepth}{0}
\setcounter{MaxMatrixCols}{20}

\title{Assignment 4}
\author{Joshua Hwang (44302650)}
\date{25 May}

\begin{document}
\maketitle

\section{ENIGMA}
\subsection{Keyspace}
Each rotor of the ENIGMA can have 26 possible starting positions
(An English letter must be chosen). Thus the rotors have $26^3$
possible states.

The plugboard can have up to 6 wires.
Consider the case with all 6 wires used. We pick the first two letters out of
${}^{26}C_2$ then ${}^{24}C_2$ and so on for each wire giving us.
\[
    {}^{26}C_2 \times {}^{24}C_2 \times {}^{22}C_2 \times {}^{20}C_2
        \times {}^{18}C_2 \times {}^{16}C_2
\]

Since there is no order to the pairs we divide by $6!$. We repeat this
process for less than 6 wires. Our results are below,
\begin{align*}
    \text{6 wires} &= 100391791500 \\
    \text{5 wires} &= 5019589575 \\
    \text{4 wires} &= 164038875 \\
    \text{3 wires} &= 3453450 \\
    \text{2 wires} &= 44850 \\
    \text{1 wire}  &= 325 \\
\end{align*}

From here we sum them up to give 105578918575 for wire combinations.
Thus the ENIGMA keyspace is $105578918575 \times 26^3 1855655072874200$.
In scientific notation this is $1.86 \times 10^15$

\subsection{Knowing rotor products}

First looking at $A$ and $D$ we can see $A$ may pair \texttt{r} wth anything
from \texttt{omwbldxvyiqe}. From this single pairing the remaining
\texttt{guthapznfcs} are given pairs. Thus $A$ has 12 total possiblities and
$D$ is dependent on whatever configuration $A$ is.

$B$ and $E$ have the same situation except there are two choices to be made.
Pairing \texttt{u} with something in \texttt{lxnqvckbjm} and pairing \texttt{d}
with something in \texttt{fz}. From these two choices we have
$10 \times 2 = 20$.

Using the exact same technique shown with $BE$ we get $9 \times 3 = 27$ for
$CF$.

\subsection{Find rotor configurations}
We know \texttt{qfoezw} must be 6 of the same letter. For the first encrypted
letter we have \texttt{q} which must have come from the letters
\texttt{rguthapznfcs}. The second letter \texttt{f} must have come from
\texttt{dr} and since the encrypted letter \texttt{q} and \texttt{f}
are the same plain text we can determine that the letter must be \texttt{r}.

Just from this pairing we can determine that $A$ maps \texttt{r} to \texttt{q}.
Now the rotors can be deduced,
\begin{verbatim}
-----------------------------------For A---------------------------------------
(r, q)
(g, i)
(u, y)
(t, v)
(h, x)
(a, d)
(p, l)
(z, b)
(n, w)
(f, m)
(c, o)
(s, e)

(j, k)

-----------------------------------For D---------------------------------------
(q, g)
(i, u)
(y, t)
(v, h)
(x, a)
(d, p)
(l, z)
(b, n)
(w, f)
(m, c)
(o, s)
(e, r)

(j, k)
\end{verbatim}

Now we can easily determine that \texttt{h} maps to \texttt{x} for rotor $A$.
So we also know $B$ will map \texttt{x} to \texttt{o}.
We know $B$ maps \texttt{r} to \texttt{f} which allows us to determine
\texttt{(r, f)} and \texttt{(d, z)} pairs.

We also know \texttt{x} to \texttt{o} which gives us complete knowledge of
rotors $B$ and $E$.
\begin{verbatim}
-----------------------------------For B---------------------------------------
(r, f)
(d, z)

(o, x)
(w, l)
(a, m)
(h, j)
(t, b)
(g, k)
(p, c)
(u, v)
(i, q)
(y, n)

(e, s)

-----------------------------------For E---------------------------------------
(f, d)
(z, r)

(x, w)
(l, a)
(m, h)
(j, t)
(b, g)
(k, p)
(c, u)
(v, i)
(q, y)
(n, o)

(e, s)
\end{verbatim}

The process is repeated again for $CF$.
\begin{verbatim}
-----------------------------------For C---------------------------------------
(x, m)
(y, h)
(b, s)

(o, r)
(w, v)
(a, c)
(e, l)
(u, z)
(d, k)
(j, n)
(i, g)
(p, t)

(q, f)

-----------------------------------For F---------------------------------------
(m, y)
(h, b)
(s, x)

(r, w)
(v, a)
(c, e)
(l, u)
(z, d)
(k, j)
(n, i)
(g, p)
(t, o)

(q, f)
\end{verbatim}

\section{Substitution-permutation network}
The operation to create a differential table is quite straightforward,
\begin{align*}
    x^* &= x \oplus \Delta x \\
    y^* &= S(x^*) \\
    \Delta y &= y^* \oplus y \\
\end{align*}

Where the function $S(x)$ is the S-box.
The above process constructed the table below,
\begin{verbatim}
           |          delta Y          |
  X  |  Y  | delta X = 1 | delta X = 9 |
-----+-----+-------------+-------------+
  0  |  e  |      c      |      a      |
  1  |  2  |      c      |      d      |
  2  |  1  |      2      |      b      |
  3  |  3  |      2      |      6      |
  4  |  d  |      4      |      1      |
  5  |  9  |      4      |      1      |
  6  |  0  |      6      |      b      |
  7  |  6  |      6      |      1      |
  8  |  f  |      b      |      d      |
  9  |  4  |      b      |      a      |
  a  |  5  |      f      |      6      |
  b  |  9  |      f      |      b      |
  c  |  8  |      4      |      1      |
  d  |  c  |      4      |      1      |
  e  |  7  |      c      |      1      |
  f  |  b  |      c      |      b      |

\end{verbatim}

ZZZ also convert to a frequency table (number of occurences)
We have also done a frequency table
\begin{verbatim}
      |                          delta Y                              |
 d    | 0 | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 | 9 | a | b | c | d | e | f |
 e ---+---+---+---+---+---+---+---+---+---+---+---+---+---+---+---+---+
 l  1 | 0 | 0 | 2 | 0 | 4 | 0 | 2 | 0 | 0 | 0 | 0 | 2 | 4 | 0 | 0 | 2 |
 t ---+---+---+---+---+---+---+---+---+---+---+---+---+---+---+---+---+
 a  9 | 0 | 6 | 0 | 0 | 0 | 0 | 2 | 0 | 0 | 0 | 2 | 4 | 0 | 2 | 0 | 0 |

 X
\end{verbatim}

\subsection{Find specific differential trail}
We attempt some rough working to deduce some facts about our search.
The plan is to split $\frac{27}{2048}$ into a set of probabilities that
are present in our table above.

Since we're looking for a probablity of $\frac{27}{2048}$ and all our
probabilities are with a denominator of 16 we can change $\frac{27}{2048}$
to have a denominator as a power of 16. We know $16 = 2^4$ and $2048 = 2^11$.
Thus we attempt to multiply our fraction by $2^1$ first.
\[
    \frac{27}{2048} = \frac{54}{4096} = \frac{3^2 \times 6}{16^3}
\]

But since we don't have a probability with a 3 or a 9 this cannot be the
combination of probabilities we're trying to find. We now try the next
multiple of 16 by multiplying by $2^5$.
\[
    \frac{27}{2048} = \frac{864}{65536} = \frac{6^3 \times 4}{16^4}
\]

These numbers of present in our table thus we could attempt to reach 4
S-boxes with 3 conversions from 9 to 1 and a conversion with frequency 4.
(Note: we could attempt to multiply by another 16 but that does not give
us a possible combination.)

\begin{verbatim}
1001 0000 0000 1001 input

0001 0000 0000 0001 sbox 6/16 * 6/16
0000 0000 0000 1001 permute

0000 0000 0000 0001 sbox 6/16
0000 0000 0000 0001 permute

0000 0000 0000 0010 sbox 4/16
0000 0001 0000 0000 permute
\end{verbatim}

\subsection{Determine which bits of $K_5$ can be deduced}
From the illustration above we see we can deduce the bits from
$K_{5,5} \to K_{5,8}$.


\end{document}
